Commutative character sheaves and geometric types for supercuspidal representations

### 2016-07-10 Geometric types.tex
Small improvement to the introduction

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from the datum introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu datum can be constructed directly from the datum, it is convenient to consider and intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
%
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations [appearing in Yu type data] are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} %\cite{Yu:models}
\end{quotation}
%
In this paper we follow the suggestion above by showing that Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence $(\oK^i,\orho_i)$, for $i=0, \ldots, d$, of types for supercuspidal representations of a sequence of $p$-adic groups.
The main result of \cite{Yu:models} explains how to find a smooth group scheme over $\O$ to with $ \underline{G}^i(\O)=\,^\circ K^i $, where $\O$ is the ring of integers of the $p$-adic field.
In this paper we show in Theorem~\ref{thm:geotypes} that if the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic, then each type $(\oK^i,\orho_i)$ in this sequence can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, 
where $\cs{F}^i$ is a virtual Weil sheaf on the Greenberg transform $G^i$ of $\underline{G}^i$, such that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i),
\end{equation}
where $\trFrob{\cs{F}^i}$ refers to the function on $G^i(k) = \underline{G}^i(\OK) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the virtual complex $\cs{F}^i$.
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob.}}
(\underline{G}^i, \cs{F}^i),
\end{tikzcd}
\]
%Theorem~\ref{thm:geotypes} shows how to construct the virtual Weil sheaf $\cs{F}^i$ from the Yu type datum, so that \eqref{eqn:intro1} holds.
We refer to the pairs $(\underline{G}^i, \cs{F}^i)$ as \emph{geometric types} for supercuspidal representations.

>>>>

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups can be built from the datum introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu datum can be constructed directly from the datum, it is convenient to consider and intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
%
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations [appearing in Yu type data] are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} %\cite{Yu:models}
\end{quotation}
%
In this paper we follow the suggestion above by showing that Yu type data are geometrizable, in the following sense.
A Yu type datum determines a sequence $(\oK^i,\orho_i)$, for $i=0, \ldots, d$, of types for supercuspidal representations of a sequence of $p$-adic groups.
The main result of \cite{Yu:models} explains how to find, for each $i=0, \ldots, d$, a smooth group scheme $\underline{G}^i$ over the ring $\O$ of integers of the $p$-adic field with $\underline{G}^i(\O)=\,^\circ K^i $.
In this paper we show in Theorem~\ref{thm:geotypes} that if the geometric component group of the reductive quotient of the special fibre of the group scheme $\underline{G}^0$ is cyclic, then each type $(\oK^i,\orho_i)$ in this sequence can be replaced by a pair $(\underline{G}^i, \cs{F}^i)$, 
where $\cs{F}^i$ is a virtual Weil sheaf on the Greenberg transform $G^i$ of $\underline{G}^i$, such that 
\begin{equation}\label{eqn:intro1}
\trFrob{\cs{F}^i} = \trace(\,^\circ\rho_i),
\end{equation}
where $\trFrob{\cs{F}^i}$ refers to the function on $G^i(k) = \underline{G}^i(\OK) = \,^\circ K^i$ obtained by evaluating the trace of the action of Frobenius on the virtual complex $\cs{F}^i$.
\[
\begin{tikzcd}
(\oK^i,\orho_i) \arrow[bend left, dashed]{rrr}{\text{geometrization}} &&& \arrow[bend left, dashed]{lll}{\text{trace of Frob.}}
(\underline{G}^i, \cs{F}^i),
\end{tikzcd}
\]
%Theorem~\ref{thm:geotypes} shows how to construct the virtual Weil sheaf $\cs{F}^i$ from the Yu type datum, so that \eqref{eqn:intro1} holds.
We refer to the pairs $(\underline{G}^i, \cs{F}^i)$ as \emph{geometric types} for supercuspidal representations.
